Perturbing SU($n$)-symmetric interactions

### 2019-11-19 collective_sun/sun_perturbation.tex
added appendix on converting muti-local operators to collective operators within the fully symmetric manifold

\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\Z}{\mathcal{Z}}

@@

@@

\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\cov}{cov}
\let\var\relax
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\diag}{diag}

@@

If we {\it construct} a perturbation $\V_2\p{\m w_\Delta}$ with
coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_2} or
\eqref{eq:cond_2_eig} with eigenvalue $\Delta/2$, then
\begin{align}
  H_0 \V_2\p{\m w_\Delta} \ket\psi
  = \p{E_0 + \Delta} \V_2\p{\m w_\Delta} \ket\psi,
\end{align}
which implies that $\V_2\p{\m w_\Delta}\ket\psi$ is a state of
definite energy that depends only on the coefficients $\m w_\Delta$.
As with the case of single-body perturbations, we can decompose any
matrix $\m w$ into its projections $\m w_\Delta$ onto the eigenspace
of matrices satisfying \eqref{eq:cond_2} with eigenvalue $\Delta/2$,
i.e.~$\m w=\sum_\Delta\m w_\Delta$, which allows us to expand
\begin{align}
  \V_2\p{\m w} = \sum_\Delta \V_2\p{\m w_\Delta},
\end{align}
and in turn simplify
\begin{align}
  H_2^{(2)} = - \P_0 \V_2\p{\m w} \E \V_2\p{\m w} \P_0
  = -\sum_{\Delta>0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ can be written in
terms collective operators using techniques similar to those in
Appendix \ref{sec:sym_prod}.  This calculation is partially completed
in Appendix \ref{sec:PXYXYP}.

@@

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \V_2^2 \P_0
  \EQFS \f14 \sum_{k,\ell,p,q}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q,
  \label{eq:PXYXYP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and the coefficients $\m w$ satisfy
\eqref{eq:cond_2} with eigenvalue $\Delta\p{\m w}\ne0$, which implies
that $\m w$ is mean-zero.  To simplify the sum in
\eqref{eq:PXYXYP_start}, we first consider the terms with
$p,q\in\set{k,\ell}$:
\begin{align}
  \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  \EQFS \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1},
  &&
  \v{\m w} \equiv \sum_{k,\ell} w_{k\ell} \ket{k\ell}.
\end{align}
The terms of the sum in \eqref{eq:PXYXYP_start} with
$p\in\set{k,\ell}$ and $q\notin\set{k,\ell}$ or vice versa are
\begin{align}
  \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  &\EQFS \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
  w_{k\ell} w_{k q} \p{X_0^2 Y_1 Y_2 + Y_0 X_0 X_1 Y_2}, \\
  \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  &\EQFS \sum_{\substack{k,\ell\\p\notin\set{k,\ell}}}
  w_{k\ell} w_{pk} \p{X_0 Y_0 X_1 Y_2 + Y_0^2 X_1 X_2},
\end{align}
where
\begin{align}
  \sum_{\substack{k,\ell\\p\notin\set{k,\ell}}} w_{k\ell} w_{pk}
  = \sum_{k,\ell,p} w_{k\ell} w_{kp} - \sum_{k,\ell} w_{k\ell}^2
  = \v w\c\v w - \v{\m w}\c\v{\m w},
  &&
  \v w \equiv \sum_{k,\ell} w_{k\ell} \ket{k}.
\end{align}
Finally, the terms of the sum in \eqref{eq:PXYXYP_start} with
$p,q\notin\set{k,\ell}$ are
\begin{align}
  \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  \EQFS \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq} X_0 X_1 Y_2 Y_3,
\end{align}
where, using the fact that $\m w$ is mean-zero,
\begin{align}
  \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}} w_{k\ell} w_{pq}
  = - \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}} w_{k\ell} w_{pq}
  - \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq}
  - \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{k\ell} w_{pq}
  = 2 \v{\m w}\c\v{\m w} - 4 \v w\c\v w.
\end{align}
Altogether, we thus far have that
\begin{multline}
  \P_0 \V_2^2 \P_0
  \EQFS -\p{\v w\c\v w - \f12 \v{\m w}\c\v{\m w}} X_0 X_1 Y_2 Y_3
  + \f14 \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1} \\
  + \f14 \p{\v w\c\v w - \v{\m w}\c\v{\m w}}
  \p{X_0^2 Y_1 Y_2 + X_0 Y_0 X_1 Y_2 + Y_0 X_0 X_1 Y_2 + Y_0^2 X_1 X_2}.
\end{multline}
If $X=Y=s_\z$, where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin
operator, then
\begin{align}
  \P_0 \V_2^2 \P_0
  \EQFS - \sp{\f{\v w\c\v w-\v{\m w}\c\v{\m w}/2}{N_4}} S_\z^4
  + \sp{\f{\v w\c\v w-\v{\m w}\c\v{\m w}}{4N_2}
    + \f{\p{3N-4}\p{\v w\c\v w - \v{\m w}\c\v{\m w}/2}}{2N_4}} S_\z^2,
\end{align}
where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
operator,
\begin{align}
  N_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!},
\end{align}
and we have neglected scalar terms with no physical consequence.

>>>>

\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\Z}{\mathcal{Z}}

@@

\newcommand{\PP}{\mathbb{Z}}

@@

\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\cov}{cov}
\let\var\relax
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\counts}{counts}

@@

If we {\it construct} a perturbation $\V_2\p{\m w_\Delta}$ with
coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_2_eig} with
eigenvalue $\Delta/2$, then
\begin{align}
  H_0 \V_2\p{\m w_\Delta} \ket\psi
  = \p{E_0 + \Delta} \V_2\p{\m w_\Delta} \ket\psi,
\end{align}
which implies that $\V_2\p{\m w_\Delta}\ket\psi$ is a state of
definite energy that depends only on the coefficients $\m w_\Delta$.
As with the case of single-body perturbations, we can decompose any
matrix $\m w$ into its projections $\m w_\Delta$ onto the eigenspace
of matrices satisfying \eqref{eq:cond_2_eig} with eigenvalue
$\Delta/2$, i.e.~$\m w=\sum_\Delta\m w_\Delta$, which allows us to
expand
\begin{align}
  \V_2\p{\m w} = \sum_\Delta \V_2\p{\m w_\Delta},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2\p{\m w} \E \V_2\p{\m w} \P_0
  = -\sum_{\Delta>0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ is simplified in
Appendices \ref{sec:PXYXYP} and \ref{sec:multi_to_collective}.  If
$X=Y=s_\z$, where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin
operator, then we can expand
\begin{align}
  H_2^{(2)} \EQFS W_4\p{\m h,\m w} S_\z^4 - W_2\p{\m h,\m w} S_\z^2,
\end{align}
where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
operator, $W_k\p{\m h,\m w}$ are scalar coefficients, and we have
neglected scalar terms in $H_2^{(2)}$ without physical consequence.
In order to expand the coefficients $W_k\p{\m h,\m w}$, for any matrix
$\m m$ we define the vectors
\begin{align}
  \v{\m m} \equiv \sum_{p,q} m_{pq} \ket{pq},
  &&
  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
\end{align}
using which we have
\begin{align}
  W_4\p{\m h,\m w}
  \equiv \sum_\Delta \f1\Delta \f{\v w_\Delta\c\v w_\Delta
    - \v{\m w}_\Delta\c\v{\m w}_\Delta/2}{N!_4},
\end{align}
\begin{align}
  W_2\p{\m h,\m w}
  \equiv \sum_\Delta \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
      - \v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
        - \v{\m w}_\Delta\c\v{\m w}_\Delta/2}}{2N!_4}},
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}.
\end{align}
Note that the dependence of the coefficients of $W_k\p{\m h,\m w}$ on
the interaction matrix $\m h$ is implicit in the decomposition of the
coefficient matrix $\m w$ into components $\m w_\Delta$ as
$\m w=\sum_\Delta\m w_\Delta$, where $\m w_\Delta$ is the projection
of $\m w$ onto the $\p{\Delta/2}$-eigenspace of the operator
$\tilde{\m h}$ defined in \eqref{eq:h_super_mat}.

@@

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \V_2^2 \P_0
  \EQFS \f14 \sum_{k,\ell,p,q}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q,
  \label{eq:PXYXYP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and the coefficients $\m w$ satisfy
\eqref{eq:cond_2} with eigenvalue $\Delta\p{\m w}\ne0$, which implies
that $\m w$ is mean-zero.  To simplify the sum in
\eqref{eq:PXYXYP_start}, we first consider the terms with
$p,q\in\set{k,\ell}$:
\begin{align}
  \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  \EQFS \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1},
  &&
  \v{\m w} \equiv \sum_{k,\ell} w_{k\ell} \ket{k\ell}.
\end{align}
The terms of the sum in \eqref{eq:PXYXYP_start} with
$p\in\set{k,\ell}$ and $q\notin\set{k,\ell}$ or vice versa are
\begin{align}
  \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  &\EQFS \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
  w_{k\ell} w_{k q} \p{X_0^2 Y_1 Y_2 + Y_0 X_0 X_1 Y_2}, \\
  \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  &\EQFS \sum_{\substack{k,\ell\\p\notin\set{k,\ell}}}
  w_{k\ell} w_{pk} \p{X_0 Y_0 X_1 Y_2 + Y_0^2 X_1 X_2},
\end{align}
where
\begin{align}
  \sum_{\substack{k,\ell\\p\notin\set{k,\ell}}} w_{k\ell} w_{pk}
  = \sum_{k,\ell,p} w_{k\ell} w_{kp} - \sum_{k,\ell} w_{k\ell}^2
  = \v w\c\v w - \v{\m w}\c\v{\m w},
  &&
  \v w \equiv \sum_{k,\ell} w_{k\ell} \ket{k}.
\end{align}
Finally, the terms of the sum in \eqref{eq:PXYXYP_start} with
$p,q\notin\set{k,\ell}$ are
\begin{align}
  \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
  \EQFS \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq} X_0 X_1 Y_2 Y_3,
\end{align}
where, using the fact that $\m w$ is mean-zero,
\begin{align}
  \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}} w_{k\ell} w_{pq}
  = - \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}} w_{k\ell} w_{pq}
  - \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{k\ell} w_{pq}
  - \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{k\ell} w_{pq}
  = 2 \v{\m w}\c\v{\m w} - 4 \v w\c\v w.
\end{align}
Altogether, we thus have that
\begin{multline}
  \P_0 \V_2^2 \P_0
  \EQFS -\p{\v w\c\v w - \f12 \v{\m w}\c\v{\m w}} X_0 X_1 Y_2 Y_3
  + \f14 \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1} \\
  + \f14 \p{\v w\c\v w - \v{\m w}\c\v{\m w}}
  \p{X_0^2 Y_1 Y_2 + X_0 Y_0 X_1 Y_2
    + Y_0 X_0 X_1 Y_2 + Y_0^2 X_1 X_2}.
  \label{eq:PXYXYP}
\end{multline}
In principle, our next task is to write this result in terms of
collective operators within the fully symmetric manifold; this task
turns out to be rather non-trivial in general.  We therefore discuss
the decomposition of multi-local operators (e.g.~$X_0Y_0X_1Y_2$, a
3-local operator) into a sum of collective operators within the fully
symmetric manifold in Appendix \ref{sec:multi_to_collective}.  At the
end of Appendix \ref{sec:multi_to_collective}, we provide analytical
expressions for the special cases of
\begin{enumerate*}
\item multi-local operators that are tensor powers of a single
  operator, i.e.~of the form $X^{\otimes k}$; and
\item multi-local operators of the form $X^{\otimes k}$ with $X^2=1$.
\end{enumerate*}
These cases are relevant for e.g.~the SU(2) case with Ising-type
interactions of the form $Z\otimes Z$, with $Z$ a single-spin Pauli
operator.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Collective and multi-local operators in the fully symmetric
  manifold}
\label{sec:multi_to_collective}

Here we provide a decomposition multi-local operators into sums of
collective operators (and vice versa) within the fully symmetric
manifold of $N$ spins.  Multi-local operators are operators of the
form
\begin{align}
  O_S \equiv \bigotimes_{O\in S} O,
  \label{eq:multi_op_set}
\end{align}
where $S\equiv\set{O}$ is a set of single-spin operators $O$.  The
order of tensor factors in \eqref{eq:multi_op_set} does not matter
after projecting operators onto the fully symmetric manifold, as we
always will in this section.  While the decomposition of a collective
operator into a sum of multi-local operators is unique, the opposite
is not true: there is generally no unique decomposition of a
multi-local operator into a sum of collective operators.  We can,
however, enforce a canonical decomposition by choosing a set of
collective operators that obey the same symmetries as $O_S$, namely
invariance under arbitrary permutations of the single-spin operators
$O\in S$.  To this end, for any partition\footnote{A partition
  $P\equiv\set{p}$ of $S$ consists of mutually disjoint, non-empty
  subsets $p\subset S$ with $\bigcup_{p\in P}p=S$.} $P\equiv\set{p}$
of $S$ we define the $\abs{P}$-local operator
\begin{align}
  O_P \equiv \bigotimes_{p\in P} \bigodot_{O\in p} O,
\end{align}
where $\bigodot_{x\in\X} x$ denotes a symmetrized product of all
elements in $\X$, i.e.
\begin{align}
  \bigodot_{x\in\X} x
  \equiv \f1{\abs{\X}!}
  \sum_{\substack{\t{permutations}\\\sigma~\t{of}~\X}}
  ~ \prod_{k=1}^{\abs{\X}} x_{\sigma_k},
\end{align}
with $\sigma_k$ the $k$-th element of the permutation $\sigma$ when
expressed in one-line notation; for example,
\begin{align}
  \bigodot_{j=1}^2 x_j &= x_1 \odot x_2 = \f12 \p{x_1x_2 + x_2x_1}, \\
  \bigodot_{j=1}^3 x_j
  &= \f1{3!} \p{x_1x_2x_3 + x_1x_3x_2 + x_2x_1x_3
    + x_2x_3x_1 + x_3x_1x_2 + x_3x_2x_1}.
\end{align}
The set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$
consists of all operators that
\begin{enumerate*}
\item can be built from ordinary and tensor products of all
  single-spin operators $O\in S$, and
\item are invariant under arbitrary permutations of $S$ (after
  projection onto the fully symmetric manifold).
\end{enumerate*}
For any set $S$ of single-spin operators or partition $P$ of $S$, we
can similarly define the collective operators
\begin{align}
  \O_S \equiv \bigodot_{O\in S} \col{O},
  &&
  \O_P \equiv \bigodot_{p\in P} \col{\textstyle\bigodot_{O\in p} O},
\end{align}
where we underline a single-spin operator to denote its collective
version, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}$, where $Q^{(j)}$ is an
operator that acts with the single-spin operator $Q$ on spin $j$ and
trivially on all other spins.  By construction, the set of collective
operators $\set{\O_P:\t{partitions}~P~\t{of}~S}$ is one-to-one to the
set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$,
and it is straightforward to work out that for any set $S$ of
single-spin operators,
\begin{align}
  \O_S
  \EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
  N!_{\abs{P}}\, O_P,
  &&
  N!_k \equiv \prod_{j=0}^{k-1}\p{N-j}! = \f{N!}{\p{N-k}!},
  \label{eq:collective_to_multi_full}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold.  For any multi-local operator $O_S$ indexed by a
set $S$ of single-spin operators, we can collect expansions of the
form \eqref{eq:collective_to_multi_full} for all collective operators
built from non-empty subsets of $S$ into a system of linear equations
$\L$, and in turn solve $\L$ for $O_S$.  When
$\abs{S}\in\set{1,2,3,4}$ (and, by conjecture, all $S$), by use of a
computer algebra system (namely, Mathematica) we find that
\begin{align}
  N!_{\abs{S}}\, O_S
  &\EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
  \p{-1}^{\abs{S}-\abs{P}} \O_P \times \prod_{p\in P}\p{\abs{p}-1}!
  \label{eq:multi_to_collective_full} \\
  &\EQFS \f1{\abs{S}!} \sum_{\substack{\t{permutations}\\\Pi~\t{of}~S}}
  \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\Pi}}
  \p{-1}^{\abs{S}-\abs{\tilde P}} \O_{\tilde P}
  \times \prod_{p\in\tilde P} \p{\abs{p}-1}!
  \label{eq:multi_to_collective_full_simp}
\end{align}
where we identify permutations $\Pi$ of $S$ with an ordered list
(tuple) $\p{O_1,O_2,\cdots,O_{\abs{S}}}$ of all single-spin operators
in $S$, and we enforce that partitions $\tilde P$ of an ordered list
$\Pi$, as well as parts $p\in\tilde P$, are lexicographically ordered
in accordance with the ordering of operators in $\Pi$.  If
$S=\set{X,Y,Z}$, for example, then the ordered partitions of the
trivial permutation $\Pi=\p{X,Y,Z}$ are
\begin{align}
  \p{\p{X,Y,Z}}, &&
  \p{\p{X,Y},\p{Z}}, &&
  \p{\p{X,Z},\p{Y}}, &&
  \p{\p{X},\p{Y,Z}}, &&
  \p{\p{X},\p{Y},\p{Z}}.
\end{align}
For ordered lists $\tilde S$ and ordered partitions $\tilde P$, we
define
\begin{align}
  O_{\tilde S} \equiv \bigotimes_{O\in\tilde S} O,
  &&
  O_{\tilde P} \equiv \bigotimes_{p\in\tilde P} \prod_{O\in p} O,
  &&
  \O_{\tilde S} \equiv \prod_{O\in\tilde S} \col{O},
  &&
  \O_{\tilde P}
  \equiv \prod_{p\in\tilde P} \col{\textstyle\prod_{O\in p} O},
\end{align}
in which case we can also expand
\begin{align}
  \O_{\tilde S} \EQFS
  \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\tilde S}}
  N!_{\abs{\tilde P}} O_{\tilde P},
\end{align}
in analogy to \eqref{eq:collective_to_multi_full}.

While the expansions in \eqref{eq:multi_to_collective_full} and
\eqref{eq:multi_to_collective_full_simp} closely resemble each other,
in practice the latter involves less computational overhead due to the
lack of symmetrization in the construction of $\O_{\tilde P}$.  There
are, in principle, even more economical decompositions of a
multi-local operators into collective opreators, for example
\begin{align}
  N!_2\, X\otimes Y
  &\EQFS \col{X}\,\col{Y} - \col{XY}, \\
  N!_3\, X\otimes Y \otimes Z
  &\EQFS \col{X}\,\col{Y}\,\col{Z}
  - \col{X}\,\col{YZ} - \col{Y}\,\col{XZ} - \col{Z}\,\col{XY}
  + \col{ZXY} + \col{YXZ},
\end{align}
but these decompositions generally do not preserve the manifest
symmetry of a multi-local operator $O_S$ under arbitrary permutations
of $S$, and we do not know of any general prescription to arrive at
such decompositions aside from solving systems of equations relating
collective operators to multi-local operators.

On a final note, we consider the special case that all operators
$O\in S$ are identical, for which we provide the expansions
\begin{align}
  N!_2\, X\otimes X \EQFS \p{\col{X}}^2 - \p{\col{X^2}},
  &&
  N!_3\, X^{\otimes 3} \EQFS \p{\col{X}}^3
  - 3 \col{X} \odot \col{X^2} + 2\p{\col{X^3}},
\end{align}
\begin{align}
  N!_4\, X^{\otimes 4}
  \EQFS \p{\col{X}}^4 - 4 \p{\col{X}}^2\odot\p{\col{X^2}}
  - 2 \col{X} \p{\col{X}}^2 \col{X}
  + 3 \p{\col{X^2}}^2 + 8 \col{X} \odot \p{\col{X^3}}
  - 6 \p{\col{X^4}}.
\end{align}
If furthermore $X^2=1$, as with e.g.~the Pauli operators of SU(2),
then
\begin{align}
  N!_2\, X\otimes X \EQFS \p{\col{X}}^2 - N,
  &&
  N!_3\, X^{\otimes 3} \EQFS \p{\col{X}}^3 - \p{3N-2} \col{X},
\end{align}
\begin{align}
  N!_4\, X^{\otimes 4} \EQFS \p{\col{X}}^4
  - 2\p{3N-4} \p{\col{X}}^2 + 3 N \p{N-2}.
\end{align}
